---
title: "Homework3"
author: "Semyon Nikulin"
date: "2025-02-13"
output: html_document
---

UT eid: syn228

Github link: https://github.com/SemyonNikulin/HW3.git

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(boot)
gas <- read.csv("gasprices.csv")

```

# PROBLEM 1

## Claim: Gas stations charge more if they lack direct competition in sight. 

Evidence:


```{r, echo=FALSE, message=FALSE, warning=FALSE}
avg_prices <- gas %>%
  group_by(Competitors) %>%
  summarise(Average_Price = mean(Price, na.rm = TRUE))

ggplot(avg_prices, aes(x = Competitors, y = Average_Price, fill = Competitors)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Gas Prices: With vs. Without Competitors",
       x = "Competitors Nearby",
       y = "Average Price of Gasoline ($)") +
  theme_minimal() +
  theme(legend.position = "none")

prices_with_comp <- gas %>% 
  filter(Competitors == "Y") %>% 
  pull(Price)
prices_without_comp <- gas %>% 
  filter(Competitors == "N") %>% 
  pull(Price)

mean_with_comp <- mean(prices_with_comp, na.rm = TRUE)
mean_without_comp <- mean(prices_without_comp, na.rm = TRUE)

ci_with_comp <- t.test(prices_with_comp)$conf.int
ci_without_comp <- t.test(prices_without_comp)$conf.int

price_difference <- mean_without_comp - mean_with_comp

t_test_result <- t.test(prices_without_comp, prices_with_comp, var.equal = FALSE)
p_value <- t_test_result$p.value
diff_ci <- t_test_result$conf.int
```

Estimated with a confidence interval of 95%, gas stations without direct competition charge somewhere between -$0.01 and $0.06 more than gas stations with direct competition. However, the p-value for this t.test was 0.15, meaning that his difference is not statistically significant.

Conclusion: This claim is not supported by the data, as our p-value is above 0.05

\newpage

## Claim: The richer the area, the higher the gas prices.

Evidence:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cor_test <- cor.test(gas$Income, gas$Price)

ggplot(gas, aes(x = Income, y = Price)) +
  geom_point(alpha = 0.5) +  
  geom_smooth(method = "lm", color = "red", se = TRUE) +  
  labs(title = "Gas Prices vs. Income",
       x = "Average Income of Area ($)",
       y = "Gas Price ($)") +
  theme_minimal()
```

Estimated with a confidence level of 95%, the correlation coefficient between income and gas price is somewhere between 0.22 and 0.55. Additionally, the p-value for this test was 4.12e-5, so our results are statistically significant. 

Conclusion: There seems to a moderate, positive correlation between the how rich an area is and the gas prices in that area (the claim is supported by the data)

\newpage

## Claim: Gas stations at stoplights charge more.

Evidence:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
prices_at_stoplights <- gas %>% 
  filter(Stoplight == "Y") %>% 
  pull(Price)
prices_not_at_stoplights <- gas %>% 
  filter(Stoplight == "N") %>% 
  pull(Price)
mean_at_stoplights <- mean(prices_at_stoplights, na.rm = TRUE)
mean_not_at_stoplights <- mean(prices_not_at_stoplights, na.rm = TRUE)

t_test_result <- t.test(prices_at_stoplights, prices_not_at_stoplights, var.equal = FALSE)
avg_prices <- gas %>%
  group_by(Stoplight) %>%
  summarise(Average_Price = mean(Price, na.rm = TRUE))

ggplot(avg_prices, aes(x = Stoplight, y = Average_Price, fill = Stoplight)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Gas Prices: At Stoplights vs. Not at Stoplights",
       x = "Stoplight Presence",
       y = "Average Gas Price ($)") +
  theme_minimal() +
  scale_fill_manual(values = c("Y" = "blue", "N" = "red")) +
  theme(legend.position = "none")
```

Estimated with a confidence level of 95%, gas stations at stoplights charge somewhere between -$0.04 and $0.03 more than gas stations away from stoplights. However, the p-value for this test was 0.85, so these results are not statistically significant.

Conclusion: Because 0 is in our confidence interval and the p-value is 0.85, this claim is not supported by the data.

\newpage

## Claim: Gas stations with direct highway access charge more.

Evidence:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
prices_on_highway <- gas %>% 
  filter(Highway == "Y") %>% 
  pull(Price)
prices_off_highway <- gas %>% 
  filter(Highway == "N") %>% 
  pull(Price)

mean_on_highway <- mean(prices_on_highway, na.rm = TRUE)
mean_off_highway <- mean(prices_off_highway, na.rm = TRUE)

t_test_result <- t.test(prices_on_highway, prices_off_highway, var.equal = FALSE)

avg_prices <- gas %>%
  group_by(Highway) %>%
  summarise(Average_Price = mean(Price, na.rm = TRUE))

ggplot(avg_prices, aes(x = Highway, y = Average_Price, fill = Highway)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Gas Prices: Highway vs. Non-Highway",
       x = "Highway Access",
       y = "Average Gas Price ($)") +
  theme_minimal() +
  theme(legend.position = "none")
```

Estimated at a confidence level of 95%, the difference between gas prices at gas stations with highway access and without highway access is somewhere between 0.01 and 0.08. Additionally, the p-value for this test is 0.02, meaning the results are statistically significant.

Conclusion: This claim is supported by the data because 0 is not in the confidence interval and the p-value is less than 0.05.

\newpage

## Claim: Shell charges more than all other non-Shell brands.

Evidence:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
prices_shell <- gas %>% 
  filter(Brand == "Shell") %>% 
  pull(Price)
prices_non_shell <- gas %>% 
  filter(Brand != "Shell") %>% 
  pull(Price)
mean_shell <- mean(prices_shell, na.rm = TRUE)
mean_non_shell <- mean(prices_non_shell, na.rm = TRUE)

t_test_result <- t.test(prices_shell, prices_non_shell, var.equal = FALSE)

gas$Brand_Category <- ifelse(gas$Brand == "Shell", "Shell", "Non-Shell")

avg_prices <- gas %>%
  group_by(Brand_Category) %>%
  summarise(Average_Price = mean(Price, na.rm = TRUE))

ggplot(avg_prices, aes(x = Brand_Category, y = Average_Price, fill = Brand_Category)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Gas Prices: Shell vs. Non-Shell Brands",
       x = "Brand Type",
       y = "Average Gas Price ($)") +
  theme_minimal() +
  scale_fill_manual(values = c("Shell" = "blue", "Non-Shell" = "red")) +
  theme(legend.position = "none")

```

Estimating with a confidence level of 95%, the difference between gas prices of shell brands and non-shell brands is somewhere between -$0.01 and $0.07. However, the p-value for this test is 0.1603 and there 0 is in the confidence interval, so this difference is not statistically significant

Conclusion: This claim is not supported by the data because the p-value is greater than 0.05

# PROBLEM 2

## Part A

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sclass <- read.csv("sclass.csv")
sclass_filtered <- sclass %>%
  filter(year == 2011, trim == "63 AMG")

mileage_values <- sclass_filtered$mileage

bootstrap_mean <- function(data, indices) {
  return(mean(data[indices], na.rm = TRUE))
}
bootstrap_results <- boot(mileage_values, bootstrap_mean, R = 10000)
ci <- boot.ci(bootstrap_results, type = "perc")
cat("Number of Cars:", nrow(sclass_filtered), "\n")
cat("95% Bootstrap Confidence Interval for Average Mileage:", ci$percent[4], "-", ci$percent[5], "miles\n")
```

## Part B

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_filtered <- sclass %>%
  filter(year == 2014, trim == "550")

df_filtered <- df_filtered %>%
  mutate(isBlack = grepl("black", tolower(color)))

black_cars <- as.numeric(df_filtered$isBlack)

n_bootstrap <- 10000
n_samples <- length(black_cars)

bootstrap_proportions <- numeric(n_bootstrap)

for (i in 1:n_bootstrap) {
  resampled_data <- sample(black_cars, size = n_samples, replace = TRUE)
  bootstrap_proportions[i] <- mean(resampled_data) 
}

bootstrap_proportions_sorted <- sort(bootstrap_proportions)  
ci_lower <- bootstrap_proportions_sorted[round(0.025 * n_bootstrap)]  
ci_upper <- bootstrap_proportions_sorted[round(0.975 * n_bootstrap)] 

cat("Number of Cars:", n_samples, "\n")
cat("95% Bootstrap Confidence Interval for Proportion of Black Cars:", ci_lower, "-", ci_upper, "\n")
```

\newpage

# PROBLEM 3

## Part A
```{r, echo=FALSE, message=FALSE, warning=FALSE}
nbc <- read.csv("nbc_pilotsurvey.csv")
filtered_nbc <- nbc %>% filter(Show %in% c("Living with Ed", "My Name is Earl"))

ed_happy <- filtered_nbc %>% 
  filter(Show == "Living with Ed") %>% 
  pull(Q1_Happy)
earl_happy <- filtered_nbc %>% 
  filter(Show == "My Name is Earl") %>% 
  pull(Q1_Happy)

t_test_result <- t.test(ed_happy, earl_happy, var.equal = FALSE)
```

Question: Consider the shows “Living with Ed” and “My Name is Earl.” Is there evidence that one show consistently produces a higher mean Q1_Happy response among viewers?

Approach: I used a t-test to see if there was a significant difference in happy responses between two shows.

Evidence: Estimated with a confidence level of 95%, the difference in happy responses from the two shows is somewhere between -0.1 and 0.4. The p-value was 0.24

Conclusion: Because 0 is in the confidence interval and the p-value is greater than 0.05, so there is no evidence that one show consistently produces a higher mean Q1_Happy response among viewers.

## Part B

```{r, echo=FALSE, message=FALSE, warning=FALSE}
filtered_nbc <- nbc %>% 
  filter(Show %in% c("The Biggest Loser", "The Apprentice: Los Angeles"))

biggest_loser_annoyed <- filtered_nbc %>% 
  filter(Show == "The Biggest Loser") %>% 
  pull(Q1_Annoyed)
apprentice_annoyed <- filtered_nbc %>% 
  filter(Show == "The Apprentice: Los Angeles") %>% 
  pull(Q1_Annoyed)

t_test_result <- t.test(biggest_loser_annoyed, apprentice_annoyed, conf.level = 0.95)
```

Question: Consider the shows “The Biggest Loser” and “The Apprentice: Los Angeles.” Is there evidence that one show consistently produces a higher mean Q1_Annoyed response among viewers?

Approach: I used a t-test to see if there was a significant difference in annoyed responses between two shows.

Evidence: Estimated with a confidence level of 95%, the difference in annoyed responses from the two shows is somewhere between 0.02 and 0.52. The p-value was 0.04. 

Conclusion: Becuase 0 is not in our confidence interval and the p-value is less than 0.05, there is evidence to support that "The Apprentice" produces more annoyed responses.

## Part C

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dancing_nbc <- nbc %>% 
  filter(Show == "Dancing with the Stars")
confusing_count <- sum(dancing_nbc$Q2_Confusing >= 4)

total_responses <- nrow(dancing_nbc)

p_hat <- confusing_count / total_responses

se <- sqrt((p_hat * (1 - p_hat)) / total_responses)

z_critical <- qnorm(0.975)
margin_of_error <- z_critical * se
conf_interval <- c(p_hat - margin_of_error, p_hat + margin_of_error)
```

Question: Based on this sample of respondents, what proportion of American TV watchers would we expect to give a response of 4 or greater to the “Q2_Confusing” question?

Approach: I constructed a 95% confidence interval by finding the proportion of confusing responses to total responses and then calculating the margin of error.

Results: I found the 95% confidence interval to be from 0.04 to 0.12, centered at a proportion of 0.08. 

Conclusion: I expect 95% of American TV watchers to have a proportion of between 0.04 to 0.12 of answering a 4 or a 5 to the confusion question.

\newpage

# PROBLEM 4

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ebay <- read.csv("ebay.csv")
ebay <- ebay %>% mutate(revenue_ratio = rev_after / rev_before)

treatment_group <- ebay %>% filter(adwords_pause == 1) %>% pull(revenue_ratio)
control_group <- ebay %>% filter(adwords_pause == 0) %>% pull(revenue_ratio)

obs_diff <- mean(treatment_group) - mean(control_group)

n_simulations <- 100000
boot_diffs <- replicate(n_simulations, {
  boot_treatment <- sample(treatment_group, size = length(treatment_group), replace = TRUE)
  boot_control <- sample(control_group, size = length(control_group), replace = TRUE)
  mean(boot_treatment) - mean(boot_control)
})

conf_interval <- quantile(boot_diffs, c(0.025, 0.975))

```

Question: Does the extra traffic brought to our site from paid search results—above and beyond what we’d see if we “went organic”—justify the cost of the ads themselves?

Approach: I constructed a 95% confidence interval for the difference in the revenue ratio for the treatment and control DMAs by running a Monte Carlo simulation and finding the 0.025th and 0.975th percentiles.

Results: I found the confidence interval of the difference in revenue ratio to be from -0.09 to -0.01. This means that the treatment group had a lower revenue ratio than the control group.

Conclusion: Since 0 is not in my 95% confidence interval, I conclude that there is a significant difference in revenue ratios between the treatment group and the control group. The control group, where advertising on Google AdWords continued, had a statistically higher revenue ratio than the treatment group, which paused Google AdWords.